\documentclass[a4paper]{article}

\usepackage[a4paper, margin=3cm]{geometry}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{lmodern,textcomp}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{censor}
\usepackage[colorinlistoftodos]{todonotes}


\title{Final Assignment: Autonomous Cars}

\author{Erik Wouters, Group 13}

\date{\today}

\begin{document}
\maketitle

%\begin{abstract}
%\url{https://erikwouters.stackstorage.com/s/byOpy0NoXSEp1xk}
%\begin{enumerate}
%    \item Please shortly describe the technology. (200 to 400 words)
%    \item Take a wider perspective for the chosen technology and preform PESTLE analysis. %(200 to 400 words)
%    \item Make a comparison between the chosen technology and another technology relevant %to the chosen one (the tech delta). Please elaborate why this technology is superior %then the comparison technology. (200 to 400 words)
%    \item What are (would be) barriers for entry for such technology? (200 to 400 words)
%    \item Describe a future scenario where the technology you present is in full use (or %what we refer to as a possible second phase in the sequential back casting video). %Please describe in detail the new resulting industry structure in this scenario. (200 %to 400 words)
%    \item What are the new possible business roles that might develop due to the %development of the technology? (200 to 400 words)
%    \item Who are (or would be) early adopters for this technology? (200 to 400 words)
%    \item What is (would be) the strategy for crossing the chasm? (200 to 400 words)
%\end{enumerate}
%\end{abstract}

\section{Introduction}
\label{sec:introduction}

The automotive industry started in the 1890s when the technology was ready to combine steam engine and road wagons. In the early 1900s the car sales witnessed the first boom. It had great advantages over the prevailing personal mode of transport at the time, the horse. Today's cars are more safe, durable, convenient and less costly compared to the past.

The automotive industry is currently entering into a new era of technical innovation with the introduction of autonomous cars. This transition comes with new concerns about for instance the possible decline in car ownership, governments imposing new regulations about safety systems and powertrains and a shift in the demographics of car users, and will likely bring back some of the lost features of the horse used to offer.

\section{Wider Perspective}
\label{sec:selecting}

The automotive industry, like any other industry is affected by macro environmental factors like Political, Economical, Social, Technological, Legal, and Environmental factors. The Analysis of these factors can help businesses in developing appropriate strategies in order to minimize the effects of these factors on business \cite{CAMPBELL2005123}.

\subsection{Political}
\label{sec:political}
The automotive industry is, due to its importance to the society and economy, pressured by governments and directly dependent on national and international policies regarding innovation and testing. The testing of new technologies regarding transportation on public roads requires political support to provide an adequate testing climate while minimizing the disturbance imposed on day-to-day operations and still enhancing public safety. Therefore, on all levels of government, testing policies need to be accommodated before the technology can be sufficiently tested in the public space. Policy shaping requires a long process with a lot of actors involved and it usually takes a few years. But with innovative ideas of great impact on peoples safety in the public space, this can even take longer and will require smaller progress steps and more monitoring and fine tuning along the way. This leads to a complex and time-consuming testing process for autonomous car testing, with a slow start for the first-movers and a faster potentially less complex process when a complete set of policies have been approved.

\subsection{Economic}
\label{sec:economic}
Economic prosperity and recessions have a direct influence on the automotive industry and therefore also on the research and development budget for autonomous cars. It is likely that in developed countries the number of cars will stagnate or shrink due to new ownership models like car sharing and mobility as a service \cite{Sivak2008}. The global car sales are predicted to stagnate even in emerging markets in developing countries \cite{Sivak2008}.

\subsection{Social}
\label{sec:social}
The society has a great influence on the progress of autonomous cars by the perceived safety, need, and impact of these vehicles. The market growth is determined by the perception of societal groups and businesses, investors and subsidizers or shareholders on the added value of the new technology. A success, but also failure, can be shared with the world in a split second. Therefore, access to the public is great, but comes also with a need for precautions and need for assurance of delivering all round high quality autonomous cars. 
An important social game changer is the possibility of new car ownership models.

\subsection{Technological}
\label{sec:technological}
Technological factors influencing the uptake of autonomous cars are the speed of sensors and processors, the intelligence of the car, the level of autonomy \cite{KAUR201887}, and the usability of the car/mobility service. Its success is also dependent on the saturation level of autonomous cars in the fleet of all road transport.

\subsection{Legal}
\label{sec:not-gal}
Since the technology is so disruptive, the legal side of who is accountable in case of an accident is still to be determined. The current status is that someone eligible to drive needs to be sitting behind the wheel and needs to put their hands on the wheel at all times and is responsible for the vehicles' course of actions. With full autonomy in place, this accountability might partly shift to the manufacturer. At least at the beginning of the introduction of autonomous cars, there will be complex play of legal rights between driver, car owner, insurance companies, manufacturers and others involved about the accountability in certain situations.

\subsection{Environmental}
\label{sec:environmental}

The 7th Environment Action Programme from the European Environment Agency (EEA) includes the objective to reduce the environmental impact of mobility. Improvements of vehicle efficiency have sustained a stop on the trend of increasing greenhouse gas emissions of transportation, nevertheless a decline is yet to be established \cite{EEA}. Potentially autonomous cars, when electric, can help to shift this trend to a declining one. This is because of platooning, mobility becoming a service and smart infrastructures and services (aiming to reduce the number of trips and the travel times and optimize the vehicle fleet).

\section{Comparing measurable parameters}
\label{sec:comparing}

For the comparison, a set measurable parameters are chosen. In the following list the parameters are scored on a Likert scale \cite{likert}. ($++$) indicates that autonomous cars perform much better in this category, whereas ($--$) indicates a much worse performance than the existing manually operated cars.

\begin{enumerate}
\item Travel time [h]
    \begin{enumerate}
    \item[($+$)] Although the speed of the vehicle is not likely to change through this new technology, it may in many cases reduce the total travel time of a trip considerably. Factors contributing to shorter travel time include reduced congestion at intersections \cite{IROS12-chiu}, higher capacity of highways \cite{GONG201825} and the fact you don't have to park the car, it can drive off to the next customer.
    \end{enumerate}
\item Operating cost [€]
    \begin{enumerate}
    \item[($+$)] The operating costs of an autonomous car will likely be lower because of multiple factors. The electrical drive train requires less energy and less maintenance. The occupancy of the autonomous car can be much higher compared to cars that drive for only a few trips per day, leading to lower write-off costs. Lastly, a lower number of collisions reduces the price of insurance.
    \end{enumerate}
\item Amount of human effort needed [\% of time spent interacting with the car controls]
    \begin{enumerate}
    \item[($++$)] The autonomous car is aimed to have minimal to zero human effort needed for driving. Compared to the original cars, this is a huge difference.
    \end{enumerate}
\item Safety [$\star$]
    \begin{enumerate}
    \item[($+$)] The autonomous car, although in its development, is aimed to become safer than the standard vehicle due to the minimization of human involvement at the wheel. However, this is a topic of debate as currently the technology is still developing and not proven safer than standard cars in all traffic environments like not in high density active traffic and low speed environments. In some situations the autonomous car is safer because of its ability to calculate speed and trajectory of other vehicles, pedestrians and cyclists at an intersection or high speed road and as such avoiding collision \cite{FU2018188}.
    \end{enumerate}
\item Noise production [dB]
    \begin{enumerate}
    \item[($+$)] Autonomous cars will likely have an electrical powertrain, therefore the vehicle noise emission will be lower than that of a standard car. On top of that, the need for braking will be reduced at highways and intersection thanks to platooning and optimized intersections. This results in lower overall braking noise emission from autonomous cars.
    \end{enumerate}
\end{enumerate}

\section{Technology Barriers}
\label{sec:technology-barriers}

At the current time there are many technology barriers that inhibit autonomous cars from being widely used.

The Artificial Intelligence (AI) needs to be sufficiently advanced to drive the cars. A good metric for this is when it is possible for an AI to drive a rally car or an F1 car faster than a human driver. This would also help to gain public attention and support.

There exist concerns about the loss of jobs when people do not have to drive anymore. While this is a valid concern, I personally always compare this to the invention of the washing machine, the public back then was under the impression that this would cause all women to spend their time on the couch, because they used to spend a lot of time washing clothes, and it turned out they did not end up like that, but were able to lead a richer life as a result of this invention.

Autonomous cars need to be produced by many manufacturers. Our fleet of cars consists of many cars, so if we want to replace them with autonomous cars there need to be a sufficient amount of production facilities.

Legislation needs to be in place about autonomous cars.
Insurance needs to accept liability for autonomous car accidents. Road authorities and legal authorities need to put a legal framework in place in case of accidents or damage as a result of autonomous car crashes. Car insurance companies and private car owners will need to be fully aware of their respective responsibilities, liability and rights in case of accidents with autonomous cars. Because the technology is still in development, these legal frameworks are yet to be established. Also, they currently do not allow for the usage of autonomous cars to their full ability. Therefore the current legal framework forms an entry barrier to the autonomous car adoption.

%\begin{enumerate}
%    \item Trust
%        \item The safety of autonomous cars is still dependent %on the traffic environment. The more degrees of %freedom of other road users, the less autonomous cars %perform. 
%        \item With this technology developing rapidly, the %entry barrier is to when the industry and authorities %and potential buyers tend to feel this system is fully %safe to use.
% \end{enumerate}
%
%Autonomous cars need to be much safer than human drivers.
%Partial autonomous systems need to be widely accepted and %used.

\section{Future Scenario}
\label{sec:future-scenario}

Everybody will drive autonomous cars because it is the cheapest and most convenient way to get around. People can be productive in the car or enjoy the car entertainment.
Professional users of autonomous cars will own their own fleet. Many of the existing car manufacturers will have gone bankrupt or switched to fully autonomous car production only. Early adopters will change to provide the service of car transport together with manufacturing the cars. Some of them will have profited from this change by embracing it early enough and will be thriving.

The utilization of cars will sky-rocket to over $60\%$, only limited by the time spent charging and waiting at night. Manually operated cars will be fewer and fewer. Cities may ban them, highways may force them to use a specific lane.

Traffic jams will be history because of continuous optimization of demand and supply in the transport system, meaning not the individual but the system routes and schedules all transport requests. Making it possible to have transport as a service. Platooning and vehicle-to-vehicle communication systems will increase the road capacity by up to $200\%$ for highways.

For regional transportation the autonomous cars will be less in favour, since the scheduling system works in such way that the higher a need for transport, the more supply this route will get. This means that the majority of trips being commuter trips during the week will get the highest priority, a longer distance means either longer travel times for dropping other people of first on the way or means longer waiting times before a vehicle comes available. Therefore longer distance trips will probably (if not pre-scheduled) still work best with high-speed rail or flight provision. Making the car a more local transport mode next to the electric bicycle. 

This will influence the urban planning as signalized intersections might disappear and be changed for automated and optimized virtual intersections where every road user gets a slot dependent on the time of arrival and trajectory, scheduled and displayed on an individual level rather than periodically scheduled at the intersection.

The parking strategy in cities will change, since no-one individually owns a car. There will only be decentralized (outside the of city centre) parking provision for autonomous cars, very limited on-street parking and a lot of drop-off/pick-up bays will be provided instead. Less than a third of the amount of cars today will be on the road. More green can be allowed in the cities, noise nuisance will change.

Everyone can be picked-up in a car which provides them with the equal travel quality. The status and car ownership issue will disappear. Children, elderly and teenagers will all have equal access to car transport as adults have. People do not need to be able to drive a car. The licensing system will disappear.

Road users using a similar route will experience the same time reliability since the system prioritizes based on first-in-first-out principle combined with an efficient optimization of fleet and road capacity.

\section{New Business Roles}
\label{sec:new-business-roles}

Taxi and lease companies may change their current role to a role where they have a (large) fleet of autonomous cars that can be hauled by people who need a ride.

Manufacturers of AI algorithms and hardware that can run AI algorithms will play an important role in the transition to autonomous cars. They can develop technology that can become standard in this new industry, securing them a good position in the new value network.

Manufacturers of electric cars will support this development to gain an additional edge on traditional car makers.
Manufacturers of GPS systems will partner as well. The implementation of vehicle to vehicle communication systems (hardware) is something that need to be developed and undertaken and maintained. 

The optimization of the road traffic system is key in the success of the autonomous vehicle and needs to be continuously developed on a national and international level, together with car sharing platforms and new taxi servers.  

An autonomous car cleaning service needs to be implemented because the cars need to be cleaned between uses. This role might be taken on by companies that currently operate a carwash.

Out of date and old cars will need to be transported to countries that can use them or be recycled and rebuild into autonomous cars. It is probably wise to still have a fleet of normal operated cars for special circumstances, like severe weather. Also (periodic) inspections on autonomous cars need to be performed. 

There needs to be a firm policy measure to make the incremental transition possible from zero automation to fully autonomous car transportation. This is an especially interesting role for cities and governments.

\section{Early Adopters}
\label{sec:early-adopters}

The first people to buy self driving cars will be enthusiasts and innovators who drive many kilometers per year, especially on the highways. There will be new companies in the taxi industry that push the boundaries of autonomous cars.
Traditional car manufacturers who fail to get on board with this trend will try to stop this development to keep their market shares.
Tesla, Volvo, BMW will produce the cars, but companies like comma.ai will sell retro fit kits for existing car models.
Tesla will be the new industry leader.
Existing car manufacturers try to buy start ups that have the technology for autonomous cars, or act against the technology.
Electric vehicles will gain a very large market share.

\section{Crossing the Chasm}
\label{sec:crossing-the-chasm}

Currently the most pressing issues lay in the AI that needs to control autonomous cars. The developments in this field have been amazing over the recent years, however there is not yet a standard test for AI drivers used to evaluate the performance of an autonomous cars like there is for human drivers. This is one of the things that should be looked into so the innovators and enthusiasts in this field (like comma.ai) can have a clear objective.
Once this objective has been stated there need to be policy in place to allow autonomous cars on the road. Once this is in place autonomous cars can move from a limited feature set to a whole product solution. This will attract new investors and seed money, as well as existing car producing companies.

\bibliography{bibliography}
\bibliographystyle{unsrt}

\end{document}